\chapter{Softwares}

\section{The R Statistical Software}

The vast majority of the code used throughout this manuscript has been done in \textbf{R}. Most experiments can be rerun from the Github repository of the manuscirpt at \url{https://www.github.com/adimajo/manuscrit_these}.

More information about the \textbf{R} Statistical Software, RStudio, and git, which I used extensively during the PhD, can be found respectively in~\cite{}.

\subsection{The glmdisc package} \label{app2:glmdisc}

% Fonctions principales
The glmdisc package can be found on Github at \url{https://www.github.com/adimajo/glmdisc}. It consists in the \textbf{R} implementation of the \textit{glmdisc} algorithm for discretizing continuous attributes, merging factor levels and introducing sparse pairwise interactions proposed in Chapters~\ref{chap5} and~\ref{chap6}.

\paragraph{Quick installation guide}

As the package is hosted on Github, a simple installation procedure is to get the devtools package and run:

devtools::install\_github("adimajo/glmdisc, build\_vignette = TRUE)

The build\_vignette argument ensures the package's vignette is installed as well.

Behind company proxies however, devtools::install\_github might not work (contrary to install.packages if the proxy is well set up). A workaround is to get the httr package which allows to wrap the previous function call in with\_config(use\_proxy(YOUR\_PROXY\_SETTINGS),\dots).

\paragraph{Main functions}

Once installed, the \textbf{R} help and vignette detail the functioning of the package. Nevertheless, I should mention a few 


\subsection{Miscellaneous}

Apart from the glmdisc package, I produced a package named scoring for the purpose of \textit{Credit Scoring} practitioners, which contains the glmdisc package, the Reject Inference methods discussed thoroughly in Chapter~\ref{chap2} and detailed in Appendix~\ref{app1:reject}, enhances the discretization package containing, among others, the MDLP and $chi^2$ discretization methods to which \textit{glmdisc} is compared in Chapter~\ref{chap4}, and and the model to perform automatic segmentation discussed in Chapter~\ref{chap6}. The scoring package can be found at \url{https://www.github.com/adimajo/scoring}.

The figures that were generated by the combined used of \textbf{R} code and the tikzDevice package can be rerun and are located in the R\_CODE\_FIGURES folder of the repository.

\section{The Python programming language}

Some experiments were performed in Python, both to benefit from implementations not available in \textbf{R} and to learn this rapidly-growing multi-purpose language which machine learning packages have ``catched up'' on the exhaustivity of the \textbf{R} framework.

\subsection{The glmdisc package}

The \textit{glmdisc}-SEM algorithm is available in Python, though in inferior state of development in comparison to the \textbf{R} implementation, at the following link: \url{https://www.github.com/adimajo/glmdisc_python}.

\paragraph{Quick installation guide}

As the package is hosted on Github, a simple installation procedure is to use pip.

pip install --upgrade https://github.com/adimajo/glmdisc\_python/archive/master.tar.gz

Again, behind company proxies, it might be useful to add the --proxy=http://username:password@server:port option.

\paragraph{Main functions}

Once installed, the Python help.

\subsection{The glmdisc-NN notebooks}

As mentioned in Chapter~\ref{chap4}, the implementation of \textit{glmdisc}-NN is straightforward in terms of neural network architecture. Therefore, all experiments involving \textit{glmdisc}-NN were performed in Jupyter Notebooks. The Notebooks for experiments on simulated data can be found in the PYTHON\_NOTEBOOKS folder of the repository.
